The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging\\ with the Owens Valley Radio Observatory Long Wavelength Array

### 2018-01-18 paper.tex
begin to address the referee's report

\title{
    The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging\\
    with the Owens Valley Long Wavelength Array
}

@@

\begin{abstract}
    A host of new low-frequency radio telescopes seek to measure the 21-cm transition of neutral
    hydrogen from the early universe.  These telescopes have the potential to directly probe star
    and galaxy formation at redshifts $20 \gtrsim z \gtrsim 7$, but are limited by the dynamic range
    they can achieve against foreground sources of low-frequency radio emission. Consequently, there
    is a growing demand for modern, high-fidelity maps of the sky at frequencies below 200 MHz for
    use in foreground modeling and removal.  We describe a new widefield imaging technique for
    drift-scanning interferometers, Tikhonov-regularized $m$-mode analysis imaging.  This technique
    constructs images of the entire sky in a single synthesis imaging step with exact treatment of
    widefield effects.  We describe how the CLEAN algorithm can be adapted to deconvolve maps
    generated by $m$-mode analysis imaging. We demonstrate Tikhonov-regularized $m$-mode analysis
    imaging using the Owens Valley Long Wavelength Array (OVRO-LWA) by generating 8 new maps of the
    sky north of $\delta=-30^\circ$ with 15 arcmin angular resolution, at frequencies evenly spaced
    between 36.528~MHz and 73.152~MHz, and $\sim$800 mJy/beam thermal noise. These maps are a
    10-fold improvement in angular resolution over existing full-sky maps at comparable frequencies,
    which have angular resolutions $\ge 2^\circ$. Each map is constructed exclusively from
    interferometric observations and does not represent the globally averaged sky brightness. Future
    improvements will incorporate total power radiometry, improved thermal noise, and improved
    angular resolution -- due to the planned expansion of the OVRO-LWA to 2.6~km baselines.  These
    maps serve as a first step on the path to the use of more sophisticated foreground filters in
    21-cm cosmology incorporating the measured angular and frequency structure of all foreground
    contaminants.
\end{abstract}

@@

As a demonstration of this technique we apply Tikhonov regularized $m$-mode analysis imaging to the
Owens Valley Long Wavelength Array (OVRO-LWA), and generate a series of new low-frequency maps of
the sky between 36.528~MHz and 73.152~MHz.  These maps capture the full sky visible from the Owens
Valley Radio Observatory (OVRO) with angular resolution of $\sim 15$~arcmin.  These new maps
complement the existing full-sky maps at these frequencies with greatly improved angular resolution.

@@

The structure of this paper is as follows. In \S\ref{sec:imaging}, we present Tikhonov-regularized
$m$-mode analysis imaging, a new imaging technique that allows us to image the entire visible sky in
one coherent synthesis imaging step with exact widefield corrections. In \S\ref{sec:observations} we
describe our observations with the Owens Valley Long Wavelength Array (OVRO-LWA). In
\S\ref{sec:results} we present the sky maps and compare these maps against other low-frequency sky
maps.  In \S\ref{sec:error} we discuss some of the sources of error present in the maps, and finally
in \S\ref{sec:conclusion} we present our conclusions.

@@

The key advantage of $m$-mode analysis is the block-diagonal structure of
Equation~\ref{eq:m-mode-matrix-equation}. The computational complexity of many common matrix
operations (eg. solving a linear system of equations) is $\mathcal{O}(N^3)$.  By splitting the
equation into $N$ independent blocks, the number of floating point operations required to operate on
the full matrix is reduced by a factor of $N^2$. This computational savings is what makes this
matrix algebra approach to interferometric imaging feasible.

@@

\subsection{The Owens Valley Long Wavelength Array}

@@

The Owens Valley Long Wavelength Array (OVRO-LWA) is a 288-element interferometer located at the
Owens Valley Radio Observatory near Big Pine, California \citep{hallinan_2017}.  The OVRO-LWA is a
low-frequency instrument with instantaneous bandwidth covering 27~MHz to 85~MHz and with 24~kHz
channelization.  Each antenna stand hosts two perpendicular broadband dipoles so that there are
$288\times2$ signal paths in total. These signal paths feed into the 512-input LEDA correlator
\citep{2015JAI.....450003K}, which allows the OVRO-LWA to capture the entire visible hemisphere in a
single snapshot image.

>>>>

\title{
    The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging\\
    with the Owens Valley Radio Observatory Long Wavelength Array
}

@@

\begin{abstract}
    A host of new low-frequency radio telescopes seek to measure the 21-cm transition of neutral
    hydrogen from the early universe.  These telescopes have the potential to directly probe star
    and galaxy formation at redshifts $20 \gtrsim z \gtrsim 7$, but are limited by the dynamic range
    they can achieve against foreground sources of low-frequency radio emission. Consequently, there
    is a growing demand for modern, high-fidelity maps of the sky at frequencies below 200 MHz for
    use in foreground modeling and removal.  We describe a new widefield imaging technique for
    drift-scanning interferometers, Tikhonov-regularized $m$-mode analysis imaging.  This technique
    constructs images of the entire sky in a single synthesis imaging step with exact treatment of
    widefield effects.  We describe how the CLEAN algorithm can be adapted to deconvolve maps
    generated by $m$-mode analysis imaging. We demonstrate Tikhonov-regularized $m$-mode analysis
    imaging using the Owens Valley Radio Observatory Long Wavelength Array (OVRO-LWA) by generating
    8 new maps of the sky north of $\delta=-30^\circ$ with 15 arcmin angular resolution, at
    frequencies evenly spaced between 36.528~MHz and 73.152~MHz, and $\sim$800 mJy/beam thermal
    noise. These maps are a 10-fold improvement in angular resolution over existing full-sky maps at
    comparable frequencies, which have angular resolutions $\ge 2^\circ$. Each map is constructed
    exclusively from interferometric observations and does not represent the globally averaged sky
    brightness. Future improvements will incorporate total power radiometry, improved thermal noise,
    and improved angular resolution -- due to the planned expansion of the OVRO-LWA to 2.6~km
    baselines.  These maps serve as a first step on the path to the use of more sophisticated
    foreground filters in 21-cm cosmology incorporating the measured angular and frequency structure
    of all foreground contaminants.
\end{abstract}

@@

As a demonstration of this technique we apply Tikhonov regularized $m$-mode analysis imaging to the
Owens Valley Radio Observatory Long Wavelength Array (OVRO-LWA), and generate a series of new
low-frequency maps of the sky between 36.528~MHz and 73.152~MHz.  These maps capture the full sky
visible from the Owens Valley Radio Observatory (OVRO) with angular resolution of $\sim 15$~arcmin.
These new maps complement the existing full-sky maps at these frequencies with greatly improved
angular resolution.

@@

The structure of this paper is as follows. In \S\ref{sec:imaging}, we present Tikhonov-regularized
$m$-mode analysis imaging, a new imaging technique that allows us to image the entire visible sky in
one coherent synthesis imaging step with exact widefield corrections. In \S\ref{sec:observations} we
describe our observations with the OVRO-LWA. In \S\ref{sec:results} we present the sky maps and
compare these maps against other low-frequency sky maps.  In \S\ref{sec:error} we discuss some of
the sources of error present in the maps, and finally in \S\ref{sec:conclusion} we present our
conclusions.

@@

The key advantage of $m$-mode analysis is the block-diagonal structure of
Equation~\ref{eq:m-mode-matrix-equation}. The computational complexity of many common matrix
operations (eg. solving a linear system of equations) is $\mathcal{O}(N^3)$, where $N$ is the linear
size of the matrix.  By splitting the equation into $M$ independent blocks, the number of floating
point operations required to solve the linear system of equations is now $\mathcal{O}(N^3M^{-2})$
because each block can be manipulated independently of the other blocks. This computational savings
is what makes this matrix algebra approach to interferometric imaging feasible.

@@

\subsection{The Owens Valley Radio Observatory Long Wavelength Array}

@@

The Owens Valley Radio Observatory Long Wavelength Array (OVRO-LWA) is a 288-element interferometer
located at the Owens Valley Radio Observatory near Big Pine, California \citep{hallinan_2017}.  The
OVRO-LWA is a low-frequency instrument with instantaneous bandwidth covering 27~MHz to 85~MHz and
with 24~kHz channelization.  Each antenna stand hosts two perpendicular broadband dipoles so that
there are $288\times2$ signal paths in total. These signal paths feed into the 512-input LEDA
correlator \citep{2015JAI.....450003K}, which allows the OVRO-LWA to capture the entire visible
hemisphere in a single snapshot image.
